Patient-Specific Neural Field Modeling from Micro-Electrode Data

### 2010-07-27 ltx/data_paper/DataPaper.tex
added connectivity kernel decomp outline

Figure~\ref{fig:SpatialCrossCorrelation} shows the average (across time) cross-correlations between consecutive field observations for the different time periods of the seizure progression. By averaging over time, the effect of the disturbance is reduced allowing for the support of the kernel to be seen. The subplots of the figure show that the connectivity kernel exhibits local excitation with surround inhibition. The plots indicate the support for excitatory aspect of the kernel centred at the origin is -2~mm to 2~mm in the $x$ direction and slightly wider in the $y$ direction.

The assumption of homogeneity was tested by performing the cross-correlation analysis on 4 equal sized subsections of the field observations. The subsections consisted $6\times6$ electrode grids, such that they overlapped by one electrode row and column. The overlap was required so the subsections where large enough for the support of the kernel to be seen in the cross-correlation analysis. The results of the analysis can be seen in Figure~\ref{fig:HomogeneityTest}. The results indicate that...

\section{Estimation Algorithm}
The state-space model is given by
\begin{align}
 \mathbf x_{t+1}&=q(\mathbf x_t)\boldsymbol\theta+\xi \mathbf x_t+\boldsymbol e_t(\mathbf r) \\
 \mathbf y_t&=\mathbf C \mathbf x_t+\boldsymbol \epsilon_t,
\end{align}
where the nonlinear function, $q(\cdot)$, is defined as
\begin{equation}\label{eq:QmatrixForSigmapoints}
	q(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}'
\end{equation}
and
\begin{equation}\label{eq:DefPsi}
	\boldsymbol{\Psi}(\mathbf{r}') \triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top} (\mathbf{r}'-\mathbf{r})d\mathbf{r}},
\end{equation}
$ \boldsymbol e_t(\mathbf r)\sim \mathcal N(\mathbf 0,\boldsymbol\Sigma_e(\mathbf r,\mathbf r'))$ is temporally white and spatially colored disturbance and  $\boldsymbol\epsilon_t\sim \mathcal N(\mathbf 0,\boldsymbol\Sigma_{\epsilon})$ is a white noise process.

>>>>

Figure~\ref{fig:SpatialCrossCorrelation} shows the average (across time) cross-correlations between consecutive field observations for the different time periods of the seizure progression. By averaging over time, the effect of the disturbance is reduced allowing for the support of the kernel to be seen. The subplots of the figure show that the connectivity kernel exhibits local excitation with surround inhibition. The plots indicate the support for excitatory aspect of the kernel centred at the origin is -2~mm to 2~mm in the $x$ direction and slightly wider in the $y$ direction. Although the kernel is not isotropic, we will approximate it as being isotropic. This is for computation reasons.

The assumption of homogeneity was tested by performing the cross-correlation analysis on 4 equal sized subsections of the field observations. The subsections consisted $6\times6$ electrode grids, such that they overlapped by one electrode row and column. The overlap was required so the subsections where large enough for the support of the kernel to be seen in the cross-correlation analysis. The results of the analysis can be seen in Figure~\ref{fig:HomogeneityTest}. The results show that each subsection of the field has similar cross-correlations, with each subplot showing local excitation with surround inhibition. From this analysis, we will approximate the connectivity kernel as being homogeneous.

\subsection{Parametrisation of Connectivity Kernel}
The connectivity kernel can be decomposed as 
\begin{equation}\label{eq:DefKernelDecomp}
	 w\left(\mathbf{r},\mathbf{r}'\right) \approx \boldsymbol{\psi}^\top\left(\mathbf{r},\mathbf{r}'\right) \boldsymbol{\theta},
where $\boldsymbol{\psi}(\mathbf{r},\mathbf{r}')$ is a vector of Gaussian basis functions and $\boldsymbol{\theta}$ is a vector of scaling parameters. By assuming a Gaussian isotropic connectivity structure, the kernel basis functions can be written as $\boldsymbol{\psi}(\mathbf{r}-\mathbf{r}')$. We will assume that we know the parametric form of the connectivity basis functions from the correlation analysis, where the scaling parameters $\boldsymbol{\theta}$ are unknown.
Substituting equation~\ref{eq:DefKernelDecomp} into equation\ref{eq:ReducedForm} gives
\begin{equation}\label{eq:KernelDecompForm}
	 \int_\Omega \boldsymbol{\psi}^\top\left(\mathbf{r},\mathbf{r}'\right) \boldsymbol{\theta} f(\boldsymbol{\phi}^{\top} (\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} 
	 + \xi\mathbf{x}_t + \mathbf{e}_t(\mathbf{r}).
The connectivity kernel weights can be shifted outside the integral giving
	 \int_\Omega \boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} \boldsymbol{\theta}  
	 + \xi\mathbf{x}_t + \mathbf{e}_t(\mathbf{r}).
Now taking the basis function vector inside the inner integrand gives
	 \int_\Omega \int_\Omega \boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} \boldsymbol{\theta}  
	 + \xi\mathbf{x}_t + \mathbf{e}_t(\mathbf{r}).
Equation~\ref{eq:ReducedForm} can be simplified by exploiting the isotropy of the kernel and, hence, the basis functions used to represent it where
	\boldsymbol{\psi} (\mathbf{r}-\mathbf{r}') = \boldsymbol{\psi} (\mathbf{r}'-\mathbf{r}).
To make the simplification, we define
\begin{equation}\label{eq:DefPsi}
	\boldsymbol{\Psi}(\mathbf{r}') \triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top} (\mathbf{r}'-\mathbf{r})\textrm{d}\mathbf{r}},
which is a constant $L \times n_{\theta}$ matrix that is defined analytically (see the Supporting Information, S1, for analytic convolution of two Gaussians), where $L$ is the number of basis functions (states) and $n_{\theta}$ is the number of connectivity kernel basis functions. Substituting equation~\ref{eq:DefPsi} into~\ref{eq:ReducedForm} gives
	\mathbf{x}_{t+1} = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \boldsymbol{\theta} + \xi\mathbf{x}_t 
+ \mathbf{e}_t(\mathbf{r}).
This substitution exploits the isotropy of the connectivity kernel basis functions, allowing us to swap the convolution in equation~\ref{reduced continuous model}, which is computationally demanding, with the convolution in equation~\ref{eq:DefPsi}. This provides a dramatic increase in estimation speed, since the matrix $\boldsymbol\Psi(\mathbf{r}')$ can be calculated analytically. This is of great importance as it lowers the computational demands of the estimation algorithm.
\section{Estimation Algorithm}
The state-space model is given by
 \mathbf x_{t+1}&=q(\mathbf x_t)\boldsymbol\theta+\xi \mathbf x_t+\boldsymbol e_t(\mathbf r) \\
 \mathbf y_t&=\mathbf C \mathbf x_t+\boldsymbol \epsilon_t,
where the nonlinear function, $q(\cdot)$, is defined as
\begin{equation}\label{eq:QmatrixForSigmapoints}
	q(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}'
\begin{equation}\label{eq:DefPsi}
	\boldsymbol{\Psi}(\mathbf{r}') \triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top} (\mathbf{r}'-\mathbf{r})d\mathbf{r}},
$ \boldsymbol e_t(\mathbf r)\sim \mathcal N(\mathbf 0,\boldsymbol\Sigma_e(\mathbf r,\mathbf r'))$ is temporally white and spatially colored disturbance and  $\boldsymbol\epsilon_t\sim \mathcal N(\mathbf 0,\boldsymbol\Sigma_{\epsilon})$ is a white noise process.
